---
title: "analysis"
author: "Ari"
date: "2023-05-17"
output: html_document
---

# Load packages

```{r load-packages, message=F}
library("RSQLite")
library("gtools")
library("lubridate")
library("jsonlite")
library("rjson")
library("tidyjson")
library("jpeg")
library("egg") # for geom_custom()
library("grid")
library("brms")
library("Metrics")
library("DescTools")
library("knitr")
library("ggrepel")
library("Hmisc")
library("ggtext")
library("tidyverse")
```


```{r set-options}
opts_chunk$set(comment = "#>",
               fig.show = "hold")

theme_set(theme_classic() + 
            theme(text = element_text(size = 24)))

options(dplyr.summarise.inform = F)
```

# Experiment 2 (Speaker)

## Load Data
```{r}
con = dbConnect(SQLite(), dbname = "../../data/experiment2/speaker_experiment_anonymized.db");
df_database = dbReadTable(con, "language")
dbDisconnect(con)
```

```{r filter-to-this-study}
# filter out incompletes 
df_data = df_database %>%
  filter(status %in% 3:5,
         !str_detect(mode, "debug"),
         codeversion %in% c("forced_choice_2"))
```

## Demographics

```{r extract-demographics}
df_demographics = df_data$datastring %>% 
  spread_values(language = jstring("questiondata", "language"),
                age = jstring("questiondata", "age"),
                gender = jstring("questiondata", "gender"),
                race = jstring("questiondata", "race"),
                ethnicity = jstring("questiondata", "ethnicity"),
                feedback = jstring("questiondata", "feedback")) %>% 
  rename(participant = document.id) %>% 
  mutate(time = difftime(df_data$endhit,
                         df_data$beginhit,
                         units = "mins"))
```


### Age
```{r}
df_demographics %>%
  mutate(age = as.numeric(age)) %>%
  summarise(`Mean age` = round(mean(age, na.rm = TRUE)),
            `Age standard deviation` = sd(age, na.rm = TRUE)) 
```

### Gender
```{r gender}
df_demographics %>%
  mutate(gender = str_replace(gender, "^[fF].*", "Female"),
         gender = str_replace(gender, "^[mM].*", "Male"),
         gender = str_replace(gender, "^$", NA_character_)) %>%
  count(gender, name = "Count") %>%
  mutate(gender = recode(gender, .missing = "No response")) 
```

### Race
```{r race}

df_demographics %>% 
  mutate(race = tolower(race)) %>% 
  count(race, name = "Count")

```


### Language
```{r language}
df_demographics %>%
  mutate(language = str_replace(language, "^[eE][nN][gG].*", "English"),
         language = str_replace(language, "^$", NA_character_)) %>%
  count(language, name = "Count") %>%
  mutate(language = recode(language, .missing = "No response")) 
```

## Wrangle Response Data

```{r Extract-main-data, warning=FALSE}
df_long = df_data$datastring %>% 
  as.tbl_json() %>% 
  enter_object("data") %>%
  gather_array("order") %>% 
  enter_object("trialdata") %>% 
  gather_object("index") %>% 
  append_values_string("value") %>% 
  as_tibble() %>% 
  rename(participant = document.id) %>% 
  spread(index, value) %>% 
  rename(trial = id) %>% 
  mutate_at(vars(trial, time, play_count), ~ as.numeric(.)) %>% 
  mutate(time = time / 1000) %>% 
  select(participant, trial, description, response, time, play_count) %>% 
  arrange(participant, trial)
```

### Attention Check
```{r check-for-and-label-participants-to-exclude}
df_long = df_long %>%
  mutate(is_attention_check = trial == 30)

excluded_participants = df_long %>%
  filter(response != "no_difference" & is_attention_check) %>%
  .$participant

df_long = df_long %>%
  filter(!(participant %in% excluded_participants))

num_excluded_participants = length(excluded_participants)
```


### Aggregate Participant Responses
```{r create-dataframe-with-responses-aggregated-over-participants}
df_agg = df_long %>%
  # ignore attention check and instruction video
  filter(trial != 30) %>% 
  # aggregate over participants
  group_by(trial, response) %>%
  summarise(count = length(response)) %>%   
  ungroup() %>% 
  right_join(expand.grid(trial = seq(0,29),
                         response = c("affect",
                                      "cause",
                                      "enable",
                                      "no_difference")),
             by = c("trial", "response")) %>% 
  mutate(count = ifelse(is.na(count), 0, count),
         response = factor(response,
                           levels = c("no_difference",
                                      "affect",
                                      "enable",
                                      "cause"))) %>% 
  group_by(trial) %>% 
  mutate(data_y = count/sum(count)) %>% 
  ungroup() %>% 
  arrange(trial, response) %>% 
  select(trial, response, data_y)
```


## Load Models

### Full model and no pragmatics
```{r}

df_full_model = read.csv("../python/model_performance/speaker_full_model_predictions.csv") %>% 
  mutate(model = "full") %>% 
  select(-X)

df_sem_model = read.csv("../python/model_performance/speaker_no_pragmatics_predictions.csv") %>% 
  mutate(model = "semantics") %>% 
  select(-X)

```



```{r}
df_aspects = read.csv("aspects/aspect_dataframe.csv") %>%
  select(-X)

df_full_data = df_long %>%
  select(participant, trial, response)

```

### Regression

#### Setup Regression Noise-Aspect-Data files for crossv
```{r}

unoise_range = c(0.5, 0.6, 0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6)

df_long_simp = df_long %>% 
  select(trial, participant, response)

df_aspects_noise = df_aspects %>% 
  filter(uncertainty_noise == 0.5)

temp = left_join(df_aspects_noise,
          df_long_simp,
          by = c("trial"))

for (unoise in unoise_range) {
  
  df_aspects_noise = df_aspects %>% 
    filter(uncertainty_noise == unoise)
  
  df_aspects_noise_data = left_join(df_aspects_noise,
                                    df_long_simp,
                                    by = c("trial"))
  
  if (unoise == 1.0) {
    unoise = "1.0"
  }
  
  write.csv(df_aspects_noise_data,
            paste("aspects/noise_", unoise, ".csv", sep = ""))

}

```

#### Fit Regressions all Noises
```{r}

unoise_range = c(0.5, 0.6, 0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6)
# unoise_range = c(0.5, 0.6, 0.7)

# Comment out the fitting procedure
for (unoise in unoise_range) {
  df_aspects_noise = df_aspects %>%
  filter(uncertainty_noise == unoise)

  df_aspects_data = left_join(df_aspects_noise,
                              df_full_data,
                              by = "trial") %>%
    mutate(response = factor(response,
                             levels = c("no_difference",
                                        "affect",
                                        "enable",
                                        "cause"),
                             ordered = TRUE))

   if (unoise == 1) {
    noise_str = "1.0"
   } else {
    noise_str = unoise
  }

  filename = paste("regression/reg_noise_", noise_str,
                   sep = "")

  reg = brm(formula = response ~ whether + how + sufficient + moving + unique +
              (whether + how + sufficient + moving + unique|participant) +
              (1|trial),
            data = df_aspects_data,
            family = cumulative(probit),
            file = filename,
            seed = 1)
}

```


#### Determine best fitting regression
```{r}

reg_list = list()

unoise_range = c(0.5, 0.6, 0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6)
likelihoods = c()


i = 1
for (unoise in unoise_range) {
  
  filename = paste("regression/reg_noise_",
                   ifelse(unoise == 1.0, "1.0", unoise),
                   ".rds",
                   sep = "")
  
  reg = readRDS(filename)
  
  reg_list[[i]] = reg
  
  i = i + 1
  
}

df_reg = tibble(unoise = unoise_range,
                fit = reg_list) %>%
  mutate(likelihood = map(.x = fit,
                          .f = ~sum(log_lik(.)))) %>%
  mutate(likelihood = unlist(likelihood)) %>% 
  arrange(desc(likelihood))

```


```{r}
opt_noise = df_reg$unoise[[1]]
top_reg = df_reg$fit[[1]] 
df_aspect_noise = df_aspects %>% 
  filter(uncertainty_noise == opt_noise)
```

#### Extract Regression Predictions

```{r}

df_reg_model = as.data.frame(fitted(reg,
                              newdata = df_aspect_noise,
                              re_formula = NA)[1:30,1,]) %>% 
  rename(no_difference = 1,
         affect = 2,
         enable = 3, 
         cause = 4) %>% 
  mutate(trial = seq(0,29)) %>% 
  pivot_longer(c(no_difference,
                 affect,
                 enable,
                 cause),
               names_to = "response",
               values_to = "model_y") %>% 
  mutate(model = "regression")

write.csv(df_reg_model, "../python/model_performance/speaker_regression_predictions.csv")

```

### Combine Models
```{r}

df_top_models = rbind(df_full_model,
                      df_sem_model,
                      df_reg_model)

```

## Visualize Model Performance

### Add confidence intervals
```{r}

count_with_ci = function(df, var) {
  y = df %>%
    group_by(trial, description) %>%
    count({{var}}, name = y, .drop = FALSE)
  resampled_counts =
    map_dfr(1:100,
            function(i) {
              df %>%
                mutate(response = sample({{var}}, replace = TRUE)) %>%
                count(response, .drop = FALSE)
            }) %>%
    group_by(trial, description, response) %>%
    summarise(ymin = quantile(n, probs = 0.025),
              ymax = quantile(n, probs = 0.975))
}

df_actual =
  df_long %>%
  mutate(response = factor(response,
                           levels = c("no_difference",
                                      "affect",
                                      "enable",
                                      "cause"))) %>% 
  filter(trial != 30)

df_agg = map_dfr(1:1000,
                 function(i) {
                   df_actual %>%
                     group_by(trial) %>%
                     mutate(response = sample(response, replace = TRUE)) %>%
                     count(response, .drop = FALSE) %>%
                     mutate(n = n / sum(n))}) %>%
  group_by(trial, response) %>%
  summarise(data_ymin = quantile(n, probs = 0.025),
            data_ymax = quantile(n, probs = 0.975)) %>%
  left_join(df_actual %>%
              group_by(trial) %>%
              count(response, name = "data_y", .drop = FALSE) %>%
              mutate(data_y = data_y / sum(data_y)),
            by = c("trial", "response")) %>%
  ungroup()

write.csv(df_agg, "../../data/experiment2/participant_means.csv")
```

### Sample Cases
```{r fig.width=13, fig.height=10}
sample_cases = c(12, 18, 4, 29, 5, 7, 14, 15, 6, 9)
words = c("no_difference", "affect", "enable", "cause")
labels = c("N", "A", "E", "C")
df.plot = df_agg %>%
  filter(trial %in% sample_cases) %>%
  mutate(trial = factor(trial, levels = sample_cases),
         Clip = factor(trial, labels = 1:length(sample_cases)),
         response = factor(response,
                           levels = words,
                           labels = labels))
df.model = df_top_models %>% 
  filter(trial %in% sample_cases) %>%
  mutate(Model = model,
         trial = factor(trial, levels = sample_cases),
         Clip = factor(trial, labels = 1:length(sample_cases)),
         response = factor(response,
                           levels = words,
                           labels = labels),
         Model = factor(Model, 
                        levels = c("full", "semantics", "regression"),
                        labels = c("Full", "No Pragmatics", "No Pragmatics and No Semantics")))
func_load_image = function(clip){
  readJPEG(str_c("../../figures/trial_schematics/trial", clip, ".jpeg"))
}
df.clips = df.plot %>% 
  distinct(trial, Clip) %>% 
  arrange(Clip) %>% 
  mutate(grob = map(.x = trial, .f = ~ func_load_image(clip = .x)))

df.text = df.clips %>% 
  select(Clip) %>% 
  distinct() %>% 
  mutate(x = 0.75,
         y = 1.73)

ggplot(data = df.plot,
           mapping = aes(x = response, y = data_y)) +
  geom_bar(stat = "identity",
           fill = "lightgray",
           color = "black") +
  geom_point(data = df.model,
             mapping = aes(x = response,
                           y = model_y,
                           shape = Model,
                           fill = Model),
             size = 3,
             position = position_dodge(0.8),
             color = "black") +
  geom_errorbar(
    mapping = aes(ymin = data_ymin, ymax = data_ymax),
    width = 0
  ) +
  geom_custom(data = df.clips,
              mapping = aes(data = grob, x = 2.5, y = Inf),
              grob_fun = function(x) rasterGrob(x,
                                                interpolate = T,
                                                vjust = 0)) +
  geom_text(data = df.text,
            mapping = aes(x = x,
                          y = y,
                          label = Clip),
            hjust = 0,
            size = 11,
            color = "gray40") +
  facet_wrap(~Clip,
             nrow = 2,
             labeller = "label_both",
             scales = "free_x") +
  scale_fill_brewer(type = "qual", palette = 6, name = "") +
  scale_shape_manual(name = "",
                     values = 21:23) + 
  scale_y_continuous(breaks = seq(0, 1, 0.25), 
                     labels = str_c(seq(0, 100, 25), "%")) + 
  labs(y = "Proportion of Responses",
       caption = "N: no difference, A: affected, E: enabled, C: caused") +
  coord_cartesian(clip = "off",
                  ylim = c(0, 1)) + 
  theme(legend.text.align = 0,
        panel.grid = element_blank(),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.text = element_text(size = 16),
        plot.caption = element_text(size = 16, color = "gray20", hjust = 0.5),
        plot.margin = margin(t = 4.5, l = 0.2, r = 0.2, b = 0.1, unit = "cm"),
        panel.spacing.y = unit(5, "cm"),
        strip.background = element_blank(),
        strip.text = element_blank()) +
  guides(fill = guide_legend(override.aes = list(size = 6)))

ggsave("../../figures/paper_figures/speaker_sample_cases.pdf",
       width = 13,
       height = 10)
```


### Scatter Plots

```{r fig.width=20, fig.height=5}

best_models_all = left_join(df_top_models, df_agg, by = c("trial", "response"))

df.plot = best_models_all %>%
  rename(Model = model) %>%
  mutate(Model = factor(Model,
                        levels = c("full", "semantics", "regression"),
                        labels = c("Full Model", "No Pragmatics", "No Semantics and No Pragmatics"))) %>% 
  mutate(response = factor(response,
                           levels = c("no_difference", "affect", "enable", "cause"),
                           labels = c("no difference", "affected", "enabled", "caused")))

df.text = df.plot %>% 
  group_by(Model) %>% 
  summarise(r = cor(data_y, model_y),
            RMSE = sqrt(mean((data_y - model_y)^2))) %>% 
  ungroup() %>% 
  pivot_longer(cols = -Model) %>% 
  mutate_if(is.numeric, ~ round(., 2)) %>% 
  mutate(data_y = rep(c(0.9, 1), 3),
         model_y = 0,
         label = str_c(name, " = ", value),
         label = ifelse(name == "r", 
                        str_replace(label, "0.", "."),
                        label))

ggplot(data = df.plot,
       mapping = aes(x = model_y,
                     y = data_y)) +
  geom_smooth(method = "lm",
              se = T,
              color = "black",
              show.legend = F) +
  geom_abline(slope = 1,
              intercept = 0,
              linetype = 2) +
  geom_text(data = df.text,
            mapping = aes(label = label),
            hjust = 0,
            size = 8) + 
  geom_errorbar(mapping = aes(ymin = data_ymin,
                              ymax = data_ymax),
                alpha = 0.6,
                width = 0,
                show.legend = FALSE) +
  geom_point(mapping = aes(fill = response,
                           shape = response),
             alpha = 1.0,
             size = 4) +
  facet_grid(cols = vars(Model)) +
  scale_shape_manual(values = 21:24) + 
  scale_x_continuous(breaks = seq(0,1,0.25),
                     labels = c("0%", "25%", "50%", "75%", "100%")) +
  scale_y_continuous(breaks = seq(0,1,0.25),
                     labels = c("0%", "25%", "50%", "75%", "100%")) +
  labs(x = "Model Prediction",
       y = "Proportion of Responses") +
  theme(legend.position = c(0.27, 0.2),
        # axis.title.x = element_text(hjust = 0.12),
        panel.spacing.x = unit(1, "cm"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size=14),
        strip.text = element_text(size = 24),
        legend.box.background = element_rect(color = "black",
                                             linewidth = 1.5)) +
  guides(fill = guide_legend(title = "response: "),
         shape = guide_legend(title = "response: "))

ggsave("../../figures/paper_figures/speaker_scatter.pdf",
       width = 20,
       height = 6.5)

```

## Cross Validation

### Load Models

```{r}

df_cv_full = read.csv("../python/model_performance/cv_speaker_full_model.csv") %>% 
  mutate(model = "full") %>% 
  select(split,
         trial,
         verb,
         model_pred,
         data_val,
         use,
         model)

df_cv_sem = read.csv("../python/model_performance/cv_speaker_semantics_model.csv") %>% 
  mutate(model = "semantic") %>% 
  select(split,
         trial,
         verb,
         model_pred,
         data_val,
         use,
         model)


df_cv_reg = tibble(split = 1:100,
                   model = "regression") %>% 
  mutate(data = map(.x = split,
                    .f = ~ read.csv(sprintf("splits/split%03d.csv", .)))) %>% 
  unnest(cols = data)
```

```{r}
df_cv_reg_update = df_cv_reg %>% 
  select(split,
         trial,
         response,
         model_y,
         data_y,
         half,
         model) %>% 
  rename(verb = response,
         model_pred = model_y,
         data_val = data_y,
         use = half) %>% 
  mutate(split = split - 1)


df_cv = rbind(df_cv_full,
              df_cv_sem,
              df_cv_reg_update) %>% 
  mutate(use = factor(use,
                      levels = c("train",
                                 "test")),
         model = factor(model,
                        levels = c("full",
                                   "semantic",
                                   "regression")),
         verb = ifelse(verb == "didn't affect", "no_difference", verb))

```

### Evaluate
```{r}

df_cv_perf = df_cv %>% 
  group_by(model, split, use) %>% 
  summarise(sum_sq_err = sum((model_pred - data_val)^2),
            split_r = cor(model_pred, data_val),
            split_rmse = rmse(model_pred, data_val))

```


### Visualize

```{r fig.width=10, fig.height=3}

df_to_show = df_cv_perf %>% 
  filter(use == "test")

ggplot(df_cv_perf,
       mapping = aes(x = sum_sq_err,
                     fill = model)) +
  geom_histogram(position = "identity",
                 color = "black",
                 alpha = 0.6) +
  facet_wrap(~use,
             ncol = 2,
             scales = "free_x")
```

### Summarize r and rmse
```{r}
df_cv_perf %>% 
  group_by(use, model) %>% 
  summarise(median_r = round(median(split_r), digits=2),
            lower_r = round(quantile(split_r, probs = 0.05), digits=2),
            upper_r = round(quantile(split_r, probs = 0.95), digits=2),
            median_rmse = round(median(split_rmse), digits=2),
            lower_rmse = round(quantile(split_rmse, probs = 0.05), digits=2),
            upper_rmse = round(quantile(split_rmse, probs = 0.95), digits=2))
```

### Summarize model differences
```{r}
df_cv_perf %>% 
  pivot_wider(id_cols = c(split, use),
              names_from = model,
              values_from = c(split_r, split_rmse)) %>% 
  ungroup() %>% 
  filter(use == "test") %>% 
  mutate(diff_sem_r = split_r_full - split_r_semantic,
         diff_sem_rmse = split_rmse_semantic - split_rmse_full,
         diff_reg_r = split_r_full - split_r_regression,
         diff_reg_rmse = split_rmse_regression - split_rmse_full) %>%
  select(split,
         diff_sem_r,
         diff_sem_rmse,
         diff_reg_r,
         diff_reg_rmse) %>% 
  pivot_longer(cols = c(diff_sem_r,
                        diff_sem_rmse,
                        diff_reg_r,
                        diff_reg_rmse),
               names_to = c("model_comp", "metric"),
               names_pattern = "diff_(.*)_(.*)",
               values_to = "val") %>% 
  group_by(metric, model_comp) %>% 
  summarise(med = round(median(val), digits = 2),
            lower = round(quantile(val, probs=0.05), digits = 2),
            upper = round(quantile(val, probs=0.95), digits = 2)) %>% 
  mutate(model_comp = factor(model_comp,
                             levels = c("sem", "reg")),
         metric = factor(metric,
                         levels = c("r", "rmse")))
```

# Experiment 3 (Listener)

## Read in Data

## Load Data
```{r}
con = dbConnect(SQLite(), dbname = "../../data/experiment3/listener_experiment_anonymized.db");
df_database = dbReadTable(con, "language")
dbDisconnect(con)
```

```{r reload-experiment-2}
df_data = df_database %>%
  filter(status %in% 3:5) %>%
  filter(!str_detect(mode, "debug")) %>%
  filter(codeversion %in% c("words_to_clips_2"))
```

## Demographics

```{r extract-demographics}
df_demographics = df_data$datastring %>% 
  spread_values(language = jstring("questiondata", "language"),
                age = jstring("questiondata", "age"),
                gender = jstring("questiondata", "gender"),
                race = jstring("questiondata", "race"),
                ethnicity = jstring("questiondata", "ethnicity"),
                feedback = jstring("questiondata", "feedback")) %>% 
  rename(participant = document.id) %>% 
  mutate(time = difftime(df_data$endhit,
                         df_data$beginhit,
                         units = "mins"))
```

### Age

```{r}
df_demographics %>%
  mutate(age = as.numeric(age)) %>%
  summarise(`Mean age` = round(mean(age, na.rm = TRUE)),
            `Age standard deviation` = round(sd(age, na.rm = TRUE)))
```

### Gender

```{r gender}
df_demographics %>%
  mutate(gender = str_replace(gender, "^[fF].*", "Female"),
         gender = str_replace(gender, "^[mM].*", "Male"),
         gender = str_replace(gender, "^$", NA_character_)) %>%
  count(gender, name = "Count") %>%
  mutate(gender = recode(gender, .missing = "No response"))
```

### Race

```{r race}
df_demographics %>% 
  mutate(race = ifelse(grepl("asian", race, ignore.case = TRUE), "Asian", race),
         race = ifelse(grepl("black", race, ignore.case = TRUE), "Black", race),
         race = ifelse(grepl("caucasian", race, ignore.case = TRUE), "White", race),
         race = ifelse(grepl("native", race, ignore.case = TRUE), "Native American", race),
         race = ifelse(grepl("white", race, ignore.case = TRUE), "White", race)
         ) %>% 
  count(race, name = "Count") %>% 
  mutate(race = recode(race, .missing = "No response"))
```
### Ethnicity

```{r race}
df_demographics %>% 
  count(ethnicity, name = "Count") %>% 
  mutate(ethnicity = recode(ethnicity, .missing = "No response"),
         Count = round(Count/sum(Count), digits=2)) 
```

### Language

```{r language}
df_demographics %>%
  mutate(language = str_replace(language, "^[eE][nN][gG].*", "English"),
         language = str_replace(language, "^$", NA_character_)) %>%
  count(language, name = "Count") %>%
  mutate(language = recode(language, .missing = "No response"))
```

### Time

```{r time-taken}
df_demographics %>%
  summarise(ave_time = round(mean(time, na.rm = TRUE)),
            sd_time = round(sd(time)))
```

## Wrangle Data

### Separate Train, Test, Attenion Check

```{r load-data-long-format}
df_training_and_test = df_data$datastring %>% 
  as.tbl_json() %>% 
  enter_object("data") %>%
  gather_array("order") %>% 
  enter_object("trialdata") %>% 
  gather_object("index") %>% 
  append_values_string("value") %>% 
  as_tibble() %>% 
  rename(participant = document.id) %>% 
  spread(index, value)

df_train = df_training_and_test %>%
  filter(!is.na(training_clip_number)) %>% 
  select_if(~sum(!is.na(.)) > 0)

df_test = df_training_and_test %>% 
  filter(is.na(training_clip_number)) %>% 
  select_if(~sum(!is.na(.)) > 0) %>%
  mutate(response = as.numeric(response)/100,
         left_video = as.numeric(left_video),
         right_video = as.numeric(right_video),
         flip_vid = left_video > right_video,
         video_a = pmin(left_video, right_video),
         video_b = pmax(left_video, right_video),
         flip_response = ifelse(flip_vid, 1-response, response))

df_attention_check = df_test %>% 
  filter(str_detect(df_test$trial, "attn")) %>% 
  arrange(participant, trial) %>% 
  select(participant, trial, utterance, video_a, video_b, flip_response) %>% 
  mutate(check = ifelse(utterance == "caused", flip_response < 0.5, flip_response > 0.5)) %>% 
  group_by(participant) %>% 
  summarise(check_passed = all(check))

df_long = df_test %>%
  filter(!str_detect(df_test$trial, "attn")) %>% 
  mutate(response = flip_response,
         trial = as.numeric(trial)) %>% 
  left_join(df_attention_check, by=c("participant")) %>% 
  arrange(participant, trial) %>% 
  filter(check_passed) %>% 
  select(participant, trial, order, utterance, video_a, video_b, response)
  
```

### Video play count
```{r}

df_test %>%
  summarise(left_play_mean = round(mean(as.numeric(left_play_count)), digits=2),
            left_play_sd = round(sd(left_play_count), digits=2),
            right_play_mean = round(mean(as.numeric(right_play_count)), digits=2),
            right_play_sd = round(sd(right_play_count), digits = 2))

```

```{r}
sum(df_attention_check$check_passed)
```


### Data Descriptives

```{r}
df_long %>% 
  group_by(trial,
           utterance,
           video_a,
           video_b) %>% 
  summarise(mean_resp = round(mean(response), digits=2),
            sd_resp = round(sd(response), digits=2)) %>%
  mutate(trial = trial + 1,
         video_a = video_a + 1,
         video_b = video_b + 1)
```

### Aggregate Participant Responses

```{r participant-mean-ratings}

df_agg = df_long %>%
  group_by(trial,
           utterance,
           video_a,
           video_b) %>% 
  do(data.frame(rbind(smean.cl.boot(.$response))))

write.csv(df_agg, "../../data/experiment3/participant_means.csv")

```

## Load Models

```{r}

df_full_model = read.csv("../python/model_performance/listener_full_model_predictions.csv") %>% 
  mutate(model_version = "Full Model")

df_sem_model = read.csv("../python/model_performance/listener_no_pragmatics_predictions.csv") %>% 
  mutate(model_version = "No Pragmatics")

df_reg_model = read.csv("../python/model_performance/listener_regression_predictions.csv") %>% 
  mutate(model_version = "No Semantics and No Pragmatics")

df_model_comparison = rbind(df_full_model,
                            df_sem_model,
                            df_reg_model) %>% 
  mutate(model_version = factor(model_version,
                                levels = c("Full Model",
                                           "No Pragmatics",
                                           "No Semantics and No Pragmatics")))

```

## Visualizations


### Trial Performance
```{r fig.height=40, fig.width=6}

df_to_show =  df_long %>%
  select(participant, trial, utterance, video_a, video_b, response) %>%
  mutate(label = str_c(video_a, "-", video_b),
         trial = str_c(trial, ": ", utterance),
         trial = as.factor(trial),
         trial = factor(trial, levels = mixedsort(levels(trial))))

df_model = df_model_comparison %>% 
  mutate(model_val = video_b_prob,
         label = str_c(video_a, "-", video_b),
         trial = str_c(trial, ": ", utterance),
         trial = as.factor(trial),
         trial = factor(trial, levels = mixedsort(levels(trial)))) %>% 
  filter(model_version != "people")

df_figures =  df_to_show %>%
  filter(participant == 2) %>%
  select(trial, label, video_a, video_b) %>%
  mutate(
    video_a_fig = map(
      .x = video_a,
      .f = ~ readJPEG(str_c(
        "../../figures/trial_schematics/trial", .x, ".jpeg"
      )) %>%
        rasterGrob(interpolate = T,
                   hjust = 1)
    ),
    video_b_fig = map(
      .x = video_b,
      .f = ~ readJPEG(str_c(
        "../../figures/trial_schematics/trial", .x, ".jpeg"
      )) %>%
        rasterGrob(interpolate = T,
                   hjust = 0)
    )
  )



ggplot(data = df_to_show,
       mapping = aes(x = 1,
                     y = response)) +
  stat_summary(geom = "pointrange",
               fun.data = "mean_cl_boot") +
  geom_hline(yintercept = c(0,1),
             color = "gray60") +
  geom_hline(yintercept = c(0.5),
             color = "gray60",
             alpha = 0.5) +
  geom_point(alpha = 0.1,
             position = position_jitter(width = 0.2,
                                        height = 0)) +
  geom_point(
    data = df_model,
    mapping = aes(y = model_val,
                  fill = model_version,
                  shape = model_version),
    size = 2,
    alpha = 0.75) +
  facet_wrap(~trial,
             ncol = 1) +
  scale_fill_brewer(type = "qual", palette = 6, name = "") +
  scale_shape_manual(name = "",
                     values = 21:23) + 
  geom_custom(data = df_figures,
              mapping = aes(data = video_a_fig,
                            x = 1,
                            y = -Inf),
              grob_fun = "identity") +
  geom_custom(data = df_figures,
              mapping = aes(data = video_b_fig,
                            x = 1,
                            y = Inf),
              grob_fun = "identity") +
  geom_text(data = df_figures,
            mapping = aes(x = 1,
                          y = -0.05,
                          label = video_a)) +
  geom_text(data = df_figures,
            mapping = aes(x = 1,
                          y = 1.05,
                          label = video_b)) +
  coord_flip(clip = "off") +
  theme_void() +
  theme(panel.spacing.y = unit(1, "cm"),
        legend.position = "bottom",
        plot.margin = margin(l = 2.2,
                             r = 2.2,
                             unit = "cm"))


ggsave("figures/listener_all_trials.jpg",
       width = 6,
       height = 40)

```

### Sample Trials

```{r fig.width=8, fig.height=8}

# trial_set = c(0, 17, 22, 32, 33)
trial_set = c(16, 33, 2, 21)

descriptions = c(
  "Ball A <b>caused</b> Ball B to go through the gate.",
  "Ball A <b>enabled</b> Ball B to go through the gate.",
  "Ball A <b>affected</b> Ball B's going through the gate.",
  "Ball A <b>made no difference</b> to Ball B's going through the gate."
)

trial_label = c("A", "B", "C", "D")

df_text = tibble("trial" = trial_label,
                 "descriptions" = descriptions)


df_to_show =  df_long %>%
  select(participant, trial, utterance, video_a, video_b, response) %>%
  filter(trial %in% trial_set) %>% 
  mutate(trial = factor(trial,
                        levels = trial_set,
                        labels = trial_label),
         label = str_c(video_a, "-", video_b),
         utterance = ifelse(utterance == "no_difference",
                            "no difference",
                            utterance),
         # trial = str_c(trial, ": ", utterance),
         trial = as.factor(trial))

df_model = df_model_comparison %>% 
  filter(trial %in% trial_set) %>% 
  mutate(trial = factor(trial,
                        levels = trial_set,
                        labels = trial_label),
         label = str_c(video_a, "-", video_b),
         utterance = ifelse(utterance == "no_difference",
                            "no difference",
                            utterance),
         # trial = str_c(trial, ": ", utterance),
         trial = as.factor(trial),
         model_val = video_b_prob)

df_figures =  df_to_show %>%
  filter(participant == 2) %>%
  select(trial, label, video_a, video_b) %>%
  mutate(
    video_a_fig = map(
      .x = video_a,
      .f = ~ readJPEG(str_c(
        "../../figures/trial_schematics/trial", .x, ".jpeg"
      )) %>%
        rasterGrob(interpolate = T,
                   hjust = 1)
    ),
    video_b_fig = map(
      .x = video_b,
      .f = ~ readJPEG(str_c(
        "../../figures/trial_schematics/trial", .x, ".jpeg"
      )) %>%
        rasterGrob(interpolate = T,
                   hjust = 0)
    )
  )



ggplot(data = df_to_show,
       mapping = aes(x = 1,
                     y = response)) +
  stat_summary(geom = "pointrange",
               fun.data = "mean_cl_boot") +
  geom_hline(yintercept = c(0,1),
             color = "gray60") +
  geom_hline(yintercept = c(0.5),
             color = "gray60",
             alpha = 0.5) +
  geom_point(alpha = 0.1,
             position = position_jitter(width = 0.2,
                                        height = 0)) +
  geom_point(
    data = df_model,
    mapping = aes(y = model_val,
                  fill = model_version,
                  shape = model_version),
    size = 2,
    alpha = 0.75) +
  facet_wrap(~trial,
             ncol = 1) +
  scale_fill_brewer(type = "qual", palette = 6, name = "") +
  scale_shape_manual(name = "",
                     values = 21:23) + 
  geom_custom(data = df_figures,
              mapping = aes(data = video_a_fig,
                            x = 1,
                            y = -Inf),
              grob_fun = "identity") +
  geom_custom(data = df_figures,
              mapping = aes(data = video_b_fig,
                            x = 1,
                            y = Inf),
              grob_fun = "identity") +
  geom_text(data = df_text,
            mapping = aes(label = trial),
            x = 1.16,
            y = -0.64,
            size = 6) + 
  geom_richtext(data = df_text,
            mapping = aes(label = descriptions),
            x = 1.26,
            y = 0.5,
            size = 5,
            label.size = NA,
            fill = alpha("white", 0)) +
  coord_flip(clip = "off") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 14),
        strip.text = element_blank(),
        panel.spacing = unit(2, "line"),
        plot.margin = margin(l = 6,
                             r = 6,
                             b = 0.5,
                             t = 0.9,
                             unit = "cm"))


ggsave("../../figures/paper_figures/listener_sample_cases.pdf",
       height = 8,
       width = 8,
       device = cairo_pdf)

```

### Scatter Plot
```{r fig.width=15, fig.height=5}

df_to_show = df_agg %>% 
  left_join(df_model_comparison, by = c("trial",
                                        "utterance",
                                        "video_a",
                                        "video_b")) %>% 
  filter(model_version != "people")

df_to_show$utterance = recode_factor(df_to_show$utterance,
                                     no_difference = "no difference")

df_to_show$utterance = fct_relevel(df_to_show$utterance,
                                   "no difference",
                                   "affected",
                                   "enabled",
                                   "caused")

df_text = df_to_show %>% 
  group_by(model_version) %>% 
  summarise(r = cor(Mean, video_b_prob),
            RMSE = rmse(Mean, video_b_prob)) %>% 
  ungroup() %>% 
  pivot_longer(cols = -model_version) %>% 
  mutate_if(is.numeric, ~ round(., 2)) %>% 
  mutate(Mean = rep(c(0.9, 1), length(unique(model_version))),
         video_b_prob = 0,
         label = str_c(name, " = ", value),
         label = ifelse(name == "r", 
                        str_replace(label, "0.", "."),
                        label))

ggplot(data = df_to_show,
       mapping = aes(x=video_b_prob, y=Mean)) +
  geom_smooth(method = "lm",
              se = T,
              color = "black",
              show.legend = F) +
  geom_abline(slope = 1,
              intercept = 0,
              linetype = 2) +
  geom_errorbar(mapping = aes(ymin = Lower,
                              ymax = Upper),
                alpha = 0.6,
                width = 0,
                show.legend = F) +
  geom_text(data = df_text,
            mapping = aes(label = label),
            hjust = 0,
            size = 8) +
  geom_point(mapping = aes(fill = utterance,
                           shape = utterance),
             alpha = 1.0,
             size = 4) +
  scale_shape_manual(values = 21:24) +
  xlim(0,1) +
  ylim(0,1) +
  labs(x = "Model Prediction",
       y = "Mean Judgments") +
  facet_grid(~model_version) +
  guides(fill = guide_legend(title = "utterance: "),
         shape = guide_legend(title = "utterance: ")) +
  theme(legend.position = c(0.27,0.2),
        panel.spacing.x = unit(1, "cm"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size=14),
        strip.text = element_text(size = 20),
        legend.box.background = element_rect(color = "black",
                                             linewidth = 1.5))

ggsave("../../figures/paper_figures/listener_scatter.pdf",
       width = 15,
       height = 5)

```
## Cross Validation

```{r}

df_cv_full = read.csv("../python/model_performance/cv_listener_full.csv") %>% 
  select(-X) 

df_cv_sem = read.csv("../python/model_performance/cv_listener_no_pragmatics.csv") %>% 
  select(-X)

df_cv_reg = read.csv("../python/model_performance/cv_listener_regression.csv") %>% 
  select(-X)

df_cv = rbind(df_cv_full,
              df_cv_sem,
              df_cv_reg) %>% 
  mutate(model = factor(model,
                        levels = c("full",
                                   "no_pragmatics",
                                   "regression"),
                        labels = c("Full Model",
                                   "No Pragmatics",
                                   "No Semantics and No Pragmatics")),
         use = factor(use,
                      levels = c("train",
                                 "test"))) %>% 
  arrange(split, model, use, trial)

```



### Evaluate
```{r}

df_cv_perf = df_cv %>% 
  group_by(model, split, use) %>% 
  summarise(sum_sq_err = sum((model_pred - data_val)^2),
            split_r = cor(model_pred, data_val),
            split_rmse = rmse(model_pred, data_val),
            tr2_in_block = any((trial == 2)))

```


### Visualize

```{r fig.width=10, fig.height=3}

df_to_show = df_cv_perf %>% 
  filter(use == "test")

ggplot(df_cv_perf,
       mapping = aes(x = sum_sq_err,
                     fill = model)) +
  geom_histogram(position = "identity",
                 color = "black",
                 alpha = 0.6) +
  facet_wrap(~use,
             nrow = 2)
```

### Summarize r and rmse
```{r}
df_cv_perf %>% 
  group_by(use, model) %>% 
  summarise(median_r = round(median(split_r), digits=2),
            lower_r = round(quantile(split_r, probs = 0.05), digits=2),
            upper_r = round(quantile(split_r, probs = 0.95), digits=2),
            median_rmse = round(median(split_rmse), digits=2),
            lower_rmse = round(quantile(split_rmse, probs = 0.05), digits=2),
            upper_rmse = round(quantile(split_rmse, probs = 0.95), digits=2))
```

### Summarize model differences
```{r}
df_cv_perf %>% 
  filter(use == "test") %>% 
  mutate(model = factor(model,
                        levels = c("Full Model",
                                   "No Pragmatics",
                                   "No Semantics and No Pragmatics"),
                        labels = c("prag",
                                   "sem",
                                   "reg"))) %>% 
  pivot_wider(id_cols = split,
              names_from = "model",
              values_from = c("split_r", "split_rmse")) %>% 
  mutate(diff_r__prag_sem = split_r_prag - split_r_sem,
         diff_r__prag_reg = split_r_prag - split_r_reg,
         diff_rmse__prag_sem = split_rmse_sem - split_rmse_prag,
         diff_rmse__prag_reg = split_rmse_reg - split_rmse_prag) %>% 
  select(split, 
         diff_r__prag_sem,
         diff_r__prag_reg, 
         diff_rmse__prag_sem, 
         diff_rmse__prag_reg) %>% 
  pivot_longer(cols = -split,
               names_to = c("metric", "comp"),
               names_prefix = "diff_",
               names_sep = "__",
               values_to = "diff") %>% 
  group_by(metric,
           comp) %>% 
  summarise(med_diff = round(median(diff), digits=2),
            lower_diff = round(quantile(diff, 0.05), digits=2),
            upper_diff = round(quantile(diff, 0.95), digits=2))
```


#### View individual splits
```{r}

df_cv_split = df_cv %>% 
  filter(split == 5)

ggplot(df_cv_split,
       mapping = aes(x = model_pred,
                     y = data_val)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(rows = vars(use),
             cols = vars(model))

```

```{r}

df_cv_perf_wide = df_cv_perf %>%
  pivot_wider(id_cols = c(model, split),
              names_from = c(use),
              values_from = c(split_r, tr2_in_block)) %>% 
  mutate(test_better = split_r_test > split_r_train) 

df_cv_perf_wide %>% 
  group_by(model) %>% 
  summarise(tr2_predicts_better = sum(tr2_in_block_train == test_better),
            tr2_in_train = sum(tr2_in_block_train),
            sum_test_better = sum(test_better))

```

